---
title: "Competiting Events"
author: "Ramya Priya"
date: "March 7, 2018"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
```

Loading Bike Events
```{r}
Bike_events = readRDS("D:/527- Data Mining/TUN Data Challenge/Datasets/Bike_Events.rds")
```

### Can we quantify the effect competing events are having in our top markets? 

To Answer this question , we first analyzed our top markets.
1. top markets are taking based on active registrations. 

1. After that to see if competing events are having any impact on our markets, I first studied all the competing markets and collected the information about their cause and which markets they offer bike events. 

2. Next I collected information about our top market event registrations and fund raising.

And conclusions have been given by seeing overall analysis of registrations and funds across top 15 markets.

Data vizualizations have been created using Tableau. Please see Funds from Top Markets and Participants from Top Markets PDF files. 

```{r}
# these are top markets in  terms of active registrations pulled from Bike events
markets = c("TX","PA","MN","NY","MA","NC","CO","OH","CA","KS","MO","WA","CA","IL","NJ")

#retrieving active registrations for top markets
All.events.registrations = Bike_events%>%
  select(Security.Category.Name,Fiscal.Year,State,Active.Registrations)%>%
  group_by(Security.Category.Name,Fiscal.Year,State)%>%
  summarise(sum(Active.Registrations))%>%
  filter(State %in% markets)
write.table(All.events.registrations,"Events.registrations.State.csv", sep = ",")

```

```{r}
All.events.funds = Bike_events%>%
  select(Security.Category.Name,Fiscal.Year,Total.of.All.Confirmed.Gifts...,Total.Event.Gifts...,Total.From.Participant...,Total.Team.Gifts...,Total.Not.From.Participant...,City,State)%>%
  group_by(Security.Category.Name,Fiscal.Year,State)%>%
   filter(State %in% markets)
```

```{r}
All.events.funds$TotalAmount = rowSums(All.events.funds[,3:7])

```

```{r}
All.events.funds=All.events.funds%>%
  select(Security.Category.Name,Fiscal.Year,Total.of.All.Confirmed.Gifts...,Total.Event.Gifts...,Total.From.Participant...,Total.Team.Gifts...,Total.Not.From.Participant...,City,State,TotalAmount)%>%
  group_by(Security.Category.Name,Fiscal.Year,State)%>%
  summarise(sum(TotalAmount))
```

```{r}
write.table(All.events.funds,"All.events.funds.csv", sep = ",")
```

