---
title: "Cleaning NMSS Data"
author: "Konrad Miziolek"
date: "January 26, 2018"
output: html_document
---

Loading libraries: 
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
```

#Cleaning NMSS Data

## Cleaning donations data for 2013-2017: 

Reading in the data:

```{r}
donations2013 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2013 Bike Donations.csv")
donations2014 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2014 Bike Donations.csv")
donations2015 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2015 Bike Donations.csv")
donations2016 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2016 Bike Donations.csv")
donations2017 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2017 Bike Donations.csv")

```

Comparing column names to make sure the data is in the right order and all fields are the same: 

```{r}
cols2013 <- colnames(donations2013)
cols2014 <- colnames(donations2014)
cols2015 <- colnames(donations2015)
cols2016 <- colnames(donations2016)
cols2017 <- colnames(donations2017)


identical(cols2013, cols2014)
identical(cols2013, cols2015)
identical(cols2013, cols2016)
identical(cols2013, cols2017)


remove(cols2013, cols2014, cols2015, cols2016, cols2017)
```

Changing some columns to factors: 

```{r}
cols <- c(1, 2, 3, 4, 6, 7, 10:19, 21:32, 35, 36, 38, 39, 44:49)
donations2013[,cols] <- lapply(donations2013[,cols], factor)
donations2014[,cols] <- lapply(donations2014[,cols], factor)
donations2015[,cols] <- lapply(donations2015[,cols], factor)
donations2016[,cols] <- lapply(donations2016[,cols], factor)
donations2017[,cols] <- lapply(donations2017[,cols], factor)
```


## Refactoring certain columns: 

### Donor Gender: 

```{r}
summary(donations2013$`Donor Gender`)
levels(donations2013$`Donor Gender`) <- list(Female=c("Female", "F", "Feiale"), Male=c("Male", "M", "Iale"))
summary(donations2013$`Donor Gender`)


summary(donations2014$`Donor Gender`)
levels(donations2014$`Donor Gender`) <- list(Female=c("Female", "F", "Feiale"), Male=c("Male"))
summary(donations2014$`Donor Gender`)


summary(donations2015$`Donor Gender`)
levels(donations2015$`Donor Gender`) <- list(Female=c("Female", "F", "Femala"), Male=c("Male"))
summary(donations2015$`Donor Gender`)


summary(donations2016$`Donor Gender`)
levels(donations2016$`Donor Gender`) <- list(Female=c("Female", "F", "Femala"), Male=c("Male"))
summary(donations2016$`Donor Gender`)


summary(donations2017$`Donor Gender`)
levels(donations2017$`Donor Gender`) <- list(Female=c("Female", "F", "Femala", "Famale"), Male=c("Male", "M"))
summary(donations2017$`Donor Gender`)

```

### Prior participants:

```{r}
summary(donations2013$`Is Prior Participant`)
levels(donations2013$`Is Prior Participant`) <- list(Yes = c("Yes"))
summary(donations2013$`Is Prior Participant`)


summary(donations2014$`Is Prior Participant`)
levels(donations2014$`Is Prior Participant`) <- list(Yes = "Yes")
summary(donations2014$`Is Prior Participant`)


summary(donations2015$`Is Prior Participant`)
levels(donations2015$`Is Prior Participant`) <- list(Yes = "Yes")
summary(donations2015$`Is Prior Participant`)


summary(donations2016$`Is Prior Participant`)
levels(donations2016$`Is Prior Participant`) <- list(Yes = "Yes")
summary(donations2016$`Is Prior Participant`)


summary(donations2017$`Is Prior Participant`)
levels(donations2017$`Is Prior Participant`) <- list(Yes = "Yes")
summary(donations2017$`Is Prior Participant`)

```


### Is Team Captain

```{r}
summary(donations2013$`Is Team Captain`)
levels(donations2013$`Is Team Captain`) <- list(Yes ="TRUE", No = "FALSE")
summary(donations2013$`Is Team Captain`)


summary(donations2014$`Is Team Captain`)
levels(donations2014$`Is Team Captain`) <- list(Yes = c("TRUE", "PRUE"), No = "FALSE")
summary(donations2014$`Is Team Captain`)


summary(donations2015$`Is Team Captain`)
levels(donations2015$`Is Team Captain`) <- list(Yes = "TRUE", No = c("FALSE", "BALSE"))
summary(donations2015$`Is Team Captain`)

summary(donations2016$`Is Team Captain`)
levels(donations2016$`Is Team Captain`) <- list(Yes = c("TRUE", "PRUE"), No = "FALSE")
summary(donations2016$`Is Team Captain`)

summary(donations2017$`Is Team Captain`)
levels(donations2017$`Is Team Captain`) <- list(Yes = c("TRUE"), No = c("FALSE", "BALSE", "FALSA"))
summary(donations2017$`Is Team Captain`)

```

### Gift Payment Method

```{r}
summary(donations2013$`Gift Payment Method`)
levels(donations2013$`Gift Payment Method`) <- list(Cash = "Cash", Check= "Check", `Credit Card` = "Credit Card")
summary(donations2013$`Gift Payment Method`)


summary(donations2014$`Gift Payment Method`)
levels(donations2014$`Gift Payment Method`) <- list(Cash = "Cash", Check= "Check", `Credit Card` = "Credit Card")
summary(donations2014$`Gift Payment Method`)


summary(donations2015$`Gift Payment Method`)
# No NAs or anything to refactor

summary(donations2016$`Gift Payment Method`)
levels(donations2016$`Gift Payment Method`) <- list(Cash = "Cash", Check= "Check", `Credit Card` = c("Credit Card", "Credit Car"))
summary(donations2016$`Gift Payment Method`)


summary(donations2017$`Gift Payment Method`)
levels(donations2017$`Gift Payment Method`) <- list(Cash = "Cash", Check= "Check", `Credit Card` = c("Credit Card", "Credit Car"))
summary(donations2017$`Gift Payment Method`)


```

### Gift Type

```{r}
summary(donations2013$`Gift Type`)
levels(donations2013$`Gift Type`) <- list(Offline=c("offline", "ofbline"), Online="online")
summary(donations2013$`Gift Type`)


summary(donations2014$`Gift Type`)
levels(donations2014$`Gift Type`) <- list(Offline="offline", Online=c("online", "ojline"))
summary(donations2014$`Gift Type`)


summary(donations2015$`Gift Type`)
levels(donations2015$`Gift Type`) <- list(Offline="offline", Online="online")
summary(donations2015$`Gift Type`)


summary(donations2016$`Gift Type`)
levels(donations2016$`Gift Type`) <- list(Offline="offline", Online=c("online", "onhine", "onlina"))
summary(donations2016$`Gift Type`)
             

summary(donations2017$`Gift Type`)
levels(donations2017$`Gift Type`) <- list(Offline="offline", Online=c("online", "ojline", "onlina"))
summary(donations2017$`Gift Type`)

```

## Dealing with dates

Converting to chars: 

```{r}
#Since each dataset is the same, can identify all date fields from just 2013 data
date_cols <- which(names(donations2013)%in%c("Event Date", "Date Recorded", "Donor Opt out Date", "Registration Date", "Team Creation Date"))

donations2013[, date_cols] <- as.character(donations2013[, date_cols])
donations2014[, date_cols] <- as.character(donations2014[, date_cols])
donations2015[, date_cols] <- as.character(donations2015[, date_cols])
donations2016[, date_cols] <- as.character(donations2016[, date_cols])
donations2017[, date_cols] <- as.character(donations2017[, date_cols])


```


Splitting into date and time of day (can just get rid of the time of day for now)

```{r}

  
```


Reformatting date (R doesn't recognize dates without a leading 0, IE 05/06/2007 is valid but 5/6/2007 is not)

```{r}


```


## EDA (ignore)

### Summarizing data for EDA
```{r}
gen_2013 <- summary(donations2013$`Donor Gender`)
gen_2014 <- summary(donations2014$`Donor Gender`)
gen_2015 <- summary(donations2015$`Donor Gender`)
gen_2016 <- summary(donations2016$`Donor Gender`)
gen_2017 <- summary(donations2017$`Donor Gender`)


gender_year <- tibble(c(2013, 2014, 2015, 2016, 2017), 
                          c(gen_2013[1], gen_2014[1], gen_2015[1], gen_2016[1], gen_2017[1]), 
                          c(gen_2013[2], gen_2014[2], gen_2015[2], gen_2016[2], gen_2017[2]), 
                          c(gen_2013[3], gen_2014[3], gen_2015[3], gen_2016[3], gen_2017[3]))
                          
colnames(gender_year) <- c("Year", "Female", "Male", "NA")


year_num <- tibble(2013:2017, c(nrow(donations2013), nrow(donations2014), nrow(donations2015), nrow(donations2016), nrow(donations2017)))
colnames(year_num) <- c("Year", "Number of Participants")


tot_gifts2013 <- sum(donations2013$Gift.Amount..., na.rm = TRUE)
tot_gifts2014 <- sum(donations2014$Gift.Amount..., na.rm = TRUE)
tot_gifts2015 <- sum(donations2015$Gift.Amount..., na.rm = TRUE)
tot_gifts2016 <- sum(donations2016$Gift.Amount..., na.rm = TRUE)
tot_gifts2017 <- sum(donations2017$Gift.Amount..., na.rm = TRUE)

med_gifts2013 <- median(donations2013$Gift.Amount..., na.rm = TRUE)
med_gifts2014 <- median(donations2014$Gift.Amount..., na.rm = TRUE)
med_gifts2015 <- median(donations2015$Gift.Amount..., na.rm = TRUE)
med_gifts2016 <- median(donations2016$Gift.Amount..., na.rm = TRUE)
med_gifts2017 <- median(donations2017$Gift.Amount..., na.rm = TRUE)

tot_donations <- tibble(2013:2017,
                            c(tot_gifts2013, tot_gifts2014, tot_gifts2015, tot_gifts2016, tot_gifts2017),
                            c(med_gifts2013, med_gifts2014, med_gifts2015, med_gifts2016, med_gifts2017))



add_tot_gifts2013 <- sum(donations2013$Gift.Amount..., na.rm = TRUE)
add_tot_gifts2014 <- sum(donations2014$Gift.Amount..., na.rm = TRUE)
add_tot_gifts2015 <- sum(donations2015$Gift.Amount..., na.rm = TRUE)
add_tot_gifts2016 <- sum(donations2016$Gift.Amount..., na.rm = TRUE)
add_tot_gifts2017 <- sum(donations2017$Gift.Amount..., na.rm = TRUE)

add_med_gifts2013 <- median(donations2013$Gift.Amount..., na.rm = TRUE)
add_med_gifts2014 <- median(donations2014$Gift.Amount..., na.rm = TRUE)
add_med_gifts2015 <- median(donations2015$Gift.Amount..., na.rm = TRUE)
add_med_gifts2016 <- median(donations2016$Gift.Amount..., na.rm = TRUE)
add_med_gifts2017 <- median(donations2017$Gift.Amount..., na.rm = TRUE)

add_donations <- tibble(2013:2017,
  c(add_tot_gifts2013, add_tot_gifts2014, add_tot_gifts2015, add_tot_gifts2016, add_tot_gifts2017), 
  c(add_med_gifts2013, add_med_gifts2014, add_med_gifts2015, add_med_gifts2016, add_med_gifts2017))

```





```{r}
ggplot(data = subset(donations2013, !is.na(`Donor Gender`))) + 
  geom_bar(aes(`Donor Gender`, fill = `Donor Gender`), na.rm = TRUE) + 
  xlab("Donor Gender") + 
  ylab("Count") + 
  ggtitle("2013 NMS Participants by gender")

ggplot(data = subset(donations2013, !is.na(`Gift Payment Method`))) + 
  geom_bar(aes(`Gift Payment Method`, fill = `Donor Gender`)) + 
  xlab("Gift Payment Method") + 
  ylab("Dollars") + 
  ggtitle("2013 NMS Gift Payment Method Types by Gender")



```





