---
title: "Cleaning NMSS Data"
author: "Konrad Miziolek"
date: "January 26, 2018"
output: html_document
---

Loading libraries: 
```{r}
library(readr)
library(plyr)
library(tidyr)
library(lubridate)
library(tidyverse)
library(dplyr)
library(ggplot2)

```

#Cleaning NMSS Data

## Konrad & Ramya: Cleaning donations data for 2013-2017: 

Reading in the data:

```{r}
donations2013 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2013 Bike Donations Date Fixed.csv")
donations2014 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2014 Bike Donations Date Fixed.csv")
donations2015 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2015 Bike Donations Date Fixed.csv")
donations2016 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2016 Bike Donations Date Fixed.csv")
donations2017 <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Donations\\2017 Bike Donations Date Fixed.csv")

```

Comparing column names to make sure the data is in the right order and all fields are the same: 

```{r}
cols2013 <- colnames(donations2013)
cols2014 <- colnames(donations2014)
cols2015 <- colnames(donations2015)
cols2016 <- colnames(donations2016)
cols2017 <- colnames(donations2017)


identical(cols2013, cols2014)
identical(cols2013, cols2015)
identical(cols2013, cols2016)
identical(cols2013, cols2017)


remove(cols2013, cols2014, cols2015, cols2016, cols2017)
```

Changing some columns to factors: 

```{r}
cols <- c(1, 2, 3, 4, 6, 7, 10:19, 21:32, 35, 36, 38, 39, 44:49)
donations2013[,cols] <- lapply(donations2013[,cols], factor)
donations2014[,cols] <- lapply(donations2014[,cols], factor)
donations2015[,cols] <- lapply(donations2015[,cols], factor)
donations2016[,cols] <- lapply(donations2016[,cols], factor)
donations2017[,cols] <- lapply(donations2017[,cols], factor)
```


### Refactoring certain columns: 

#### Donor Gender: 

```{r}
#summary(donations2013$`Donor Gender`)
levels(donations2013$`Donor Gender`) <- list(Female=c("Female", "F", "Feiale"), Male=c("Male", "M", "Iale"))
#summary(donations2013$`Donor Gender`)


#summary(donations2014$`Donor Gender`)
levels(donations2014$`Donor Gender`) <- list(Female=c("Female", "F", "Feiale"), Male=c("Male"))
#summary(donations2014$`Donor Gender`)


#summary(donations2015$`Donor Gender`)
levels(donations2015$`Donor Gender`) <- list(Female=c("Female", "F", "Femala"), Male=c("Male"))
#summary(donations2015$`Donor Gender`)


#summary(donations2016$`Donor Gender`)
levels(donations2016$`Donor Gender`) <- list(Female=c("Female", "F", "Femala"), Male=c("Male"))
#summary(donations2016$`Donor Gender`)


#summary(donations2017$`Donor Gender`)
levels(donations2017$`Donor Gender`) <- list(Female=c("Female", "F", "Femala", "Famale"), Male=c("Male", "M"))
#summary(donations2017$`Donor Gender`)

```

### Prior participants:

```{r}
#summary(donations2013$`Is Prior Participant`)
levels(donations2013$`Is Prior Participant`) <- list(Yes = c("Yes"))
#summary(donations2013$`Is Prior Participant`)


#summary(donations2014$`Is Prior Participant`)
levels(donations2014$`Is Prior Participant`) <- list(Yes = "Yes")
#summary(donations2014$`Is Prior Participant`)


#summary(donations2015$`Is Prior Participant`)
levels(donations2015$`Is Prior Participant`) <- list(Yes = "Yes")
#summary(donations2015$`Is Prior Participant`)


#summary(donations2016$`Is Prior Participant`)
levels(donations2016$`Is Prior Participant`) <- list(Yes = "Yes")
#summary(donations2016$`Is Prior Participant`)


#summary(donations2017$`Is Prior Participant`)
levels(donations2017$`Is Prior Participant`) <- list(Yes = "Yes")
#summary(donations2017$`Is Prior Participant`)

```


### Is Team Captain

```{r}
#summary(donations2013$`Is Team Captain`)
levels(donations2013$`Is Team Captain`) <- list(Yes ="TRUE", No = "FALSE")
#summary(donations2013$`Is Team Captain`)


#summary(donations2014$`Is Team Captain`)
levels(donations2014$`Is Team Captain`) <- list(Yes = c("TRUE", "PRUE"), No = "FALSE")
#summary(donations2014$`Is Team Captain`)


#summary(donations2015$`Is Team Captain`)
levels(donations2015$`Is Team Captain`) <- list(Yes = "TRUE", No = c("FALSE", "BALSE"))
#summary(donations2015$`Is Team Captain`)

#summary(donations2016$`Is Team Captain`)
levels(donations2016$`Is Team Captain`) <- list(Yes = c("TRUE", "PRUE"), No = "FALSE")
#summary(donations2016$`Is Team Captain`)

#summary(donations2017$`Is Team Captain`)
levels(donations2017$`Is Team Captain`) <- list(Yes = c("TRUE"), No = c("FALSE", "BALSE", "FALSA"))
#summary(donations2017$`Is Team Captain`)

```

### Gift Payment Method

```{r}
#summary(donations2013$`Gift Payment Method`)
levels(donations2013$`Gift Payment Method`) <- list(Cash = "Cash", Check= "Check", `Credit Card` = "Credit Card")
#summary(donations2013$`Gift Payment Method`)


#summary(donations2014$`Gift Payment Method`)
levels(donations2014$`Gift Payment Method`) <- list(Cash = "Cash", Check= "Check", `Credit Card` = "Credit Card")
#summary(donations2014$`Gift Payment Method`)


#summary(donations2015$`Gift Payment Method`)
# No NAs or anything to refactor

#summary(donations2016$`Gift Payment Method`)
levels(donations2016$`Gift Payment Method`) <- list(Cash = "Cash", Check= "Check", `Credit Card` = c("Credit Card", "Credit Car"))
#summary(donations2016$`Gift Payment Method`)


#summary(donations2017$`Gift Payment Method`)
levels(donations2017$`Gift Payment Method`) <- list(Cash = "Cash", Check= "Check", `Credit Card` = c("Credit Card", "Credit Car"))
#summary(donations2017$`Gift Payment Method`)


```

### Gift Type

```{r}
#summary(donations2013$`Gift Type`)
levels(donations2013$`Gift Type`) <- list(Offline=c("offline", "ofbline"), Online="online")
#summary(donations2013$`Gift Type`)


#summary(donations2014$`Gift Type`)
levels(donations2014$`Gift Type`) <- list(Offline="offline", Online=c("online", "ojline"))
#summary(donations2014$`Gift Type`)


#summary(donations2015$`Gift Type`)
levels(donations2015$`Gift Type`) <- list(Offline="offline", Online="online")
#summary(donations2015$`Gift Type`)


#summary(donations2016$`Gift Type`)
levels(donations2016$`Gift Type`) <- list(Offline="offline", Online=c("online", "onhine", "onlina"))
#summary(donations2016$`Gift Type`)
             

#summary(donations2017$`Gift Type`)
levels(donations2017$`Gift Type`) <- list(Offline="offline", Online=c("online", "ojline", "onlina"))
#summary(donations2017$`Gift Type`)

```

### Offline status

```{r}
summary(donations2013$`Offline Status`)
summary(donations2014$`Offline Status`)
summary(donations2015$`Offline Status`)
summary(donations2016$`Offline Status`)
levels(donations2016$`Offline Status`)<- list(confirmed = c("confirmed", "confirme`"), `Teamraiser Participant Gift` = c("(Teamraiser Participant Gift", "Teamraiser Participant Gift"))
summary(donations2017$`Offline Status`)
```


### Dealing with dates

* Date format was changed in Excel * to eg 08/12/13 to enable easy conversion of dates. 

```{r}
#Since each dataset is the same, can identify all date fields from just 2013 data
date_cols <- which(names(donations2013)%in%c("Event Date", "Date Recorded", "Donor Opt-out Date", "Registration Date", "Team Creation Date"))

donations2013[, date_cols] <- lapply(donations2013[, date_cols], as.Date, "%m/%d/%y")
donations2014[, date_cols] <- lapply(donations2014[, date_cols], as.Date, "%m/%d/%y")
donations2015[, date_cols] <- lapply(donations2015[, date_cols], as.Date, "%m/%d/%y")
donations2016[, date_cols] <- lapply(donations2016[, date_cols], as.Date, "%m/%d/%y")
donations2017[, date_cols] <- lapply(donations2017[, date_cols], as.Date, "%m/%d/%y")
```

### Writing out Formatted Donations Data to an RDS file

```{r}
saveRDS(donations2013, "C:/Users/Konrad/Desktop/ETM 527/RDS Data/Donations/Donations2013.rds")
saveRDS(donations2014, "C:/Users/Konrad/Desktop/ETM 527/RDS Data/Donations/Donations2014.rds")
saveRDS(donations2015, "C:/Users/Konrad/Desktop/ETM 527/RDS Data/Donations/Donations2015.rds")
saveRDS(donations2016, "C:/Users/Konrad/Desktop/ETM 527/RDS Data/Donations/Donations2016.rds")
saveRDS(donations2017, "C:/Users/Konrad/Desktop/ETM 527/RDS Data/Donations/Donations2017.rds")


```



## Konrad: Cleaning Bike Teams Data

```{r}
teams <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Bike Teams\\2013-2017 Bike Teams Date Format Fixed.csv")
```

```{r}
#summary(teams)
factor_cols <- c(1:5, 7, 8, 10, 11, 20:22, 26)
teams[ ,factor_cols] <- lapply(teams[ ,factor_cols], factor)

# Removed time from team creation date and Event Date (in Excel)!!! 

teams$`Team Creation Date` <- as.Date(teams$`Team Creation Date`, format = '%m/%d/%y')
teams$`Event Date` <- as.Date(teams$`Event Date`, format = '%m/%d/%y')

summary(teams$`Team Division`)

levels(teams$`Team Division`) <- list(Association = "Association", Brewery="Beer/Brewery", `Bike Club`="Bike Club", `Bike Shop` = c("Bike Shop", "Bike Shops"), `Civic Team` = "Civic Team", `Club/Organization` = "Club/Organization", Corporate = c("Corporate", "Corporation"), `Family/Friends` = c("Family and Friends", "Family/Friends", "Frien`s and Family", "Friend and Family", "Friends and Family"), Ohana = "Ohana", `Ohana/Friends` = "Ohana and Friends", Open = "Open", `Open Team` = "Open Team", Organization = c("Organization", "Organization (Clubs, Civic Groups, etc.)", "Organization (Clubs; Civic Groups; etc.)", "Organization (Clubs; Civic Groups; Place of Worship; etc.)"), Other = "Other", `Place of Worship` = c("Place of Worship", "Place of worship", "Religious"), School = "School", `Small Business` = "Small Business", `Volunteer Group` = "Volunteer Group")

```

```{r}
#summary(teams$`Event Type`)
#summary(teams$`Internal Event Name`)
#summary(teams$`Event ID`)
#summary(teams$`Team Creation Date`)

#summary(teams$`Captain Email Domain`)
#summary(teams$Company)
#summary(teams$`Team Name`)
#summary(teams$`Number of Participants`)
```

```{r}
saveRDS(teams, "C:/Users/Konrad/Desktop/ETM 527/RDS Data/Bike Teams/teams.rds")
```



## Ramya: Cleaning Bike Events

```{r}
BikeEvents<-read.csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Events\\2013-2017 Bike Events.csv")
```


```{r}
str(BikeEvents)
```

```{r}
levels(BikeEvents$City)
```

```{r}
#Cleaning cities
BikeEvents$City[BikeEvents$City == 'Athens, AL ']<- "Athens"

BikeEvents$City[BikeEvents$City == 'Cherry Hill, NJ with alternate start locations']<-"Cherry Hill"
#as.factor(BikeEvents$City[BikeEvents$City == 'Ridgeland, MS']<- "Ridgeland")

BikeEvents$City[BikeEvents$City == 'St. Augustine ']<-"St. Augustine"

BikeEvents$City[BikeEvents$City == 'Sioux Falls area']<-"Sioux Falls"

BikeEvents$City[BikeEvents$City == 'Schodack ']<-"Schodack"

BikeEvents$City[BikeEvents$City == 'San Antonio ']<-"San Antonio"

BikeEvents$City[BikeEvents$City == 'New York City']<-"New York" 

BikeEvents$City[BikeEvents$City == 'New Albany ']<-"New Albany"

BikeEvents$City[BikeEvents$City == 'Keuka Park ']<-"Keuka Park"
BikeEvents$City[BikeEvents$City == 'Hollidaysburg, PA to State College']<-"Hollidaysburg" 
BikeEvents$City[BikeEvents$City == 'Duluth to White Bear Lake']<-"Duluth"
BikeEvents$City[BikeEvents$City == 'Duluth to the Twin Cities']<-"Duluth"
BikeEvents$City[BikeEvents$City == 'Fargo area']<-"Fargo"
BikeEvents$City[BikeEvents$City == 'Dallas ']<-"Dallas"
BikeEvents$City[BikeEvents$City == 'DeKalb']<-"Dekalb"
BikeEvents$City[BikeEvents$City == 'DeKalb ']<-"Dekalb"

```

```{r}
summary(BikeEvents$Event.Date)
```


```{r}
#Date Conversion in R
datecols_bikeevents<-which(names(BikeEvents)%in%c("Event.Created.Date", "Event.Date"))

for (i in datecols_bikeevents) {
  BikeEvents[,i] = as.Date(BikeEvents[,i], '%m/%d/%y')
}

```

## Saving the work
```{r}
saveRDS(BikeEvents, "C:/Users/Konrad/Desktop/ETM 527/RDS Data/Bike Events/Bike Events.rds")
```

## Jordan: Cleaning Bike Participants

### Import data
```{r}
participants <- read_csv("C:\\Users\\Konrad\\Desktop\\ETM 527\\Participants\\2013-2017 Bike MS Participants.csv")
```

### Basic data exploration
```{r}
class(participants)
dim(participants)
glimpse(participants)
head(participants, n = 5)
tail(participants, n = 5)
summary(participants)
```

Able to see that some fields need to be converted from integers to characters. Also, easily able to identify which fields will need to be rid of NA values. 

```{r}
# need to change fields from int to char 
participants$`Team ID` <- as.character(participants$`Team ID`)
class(participants$`Team ID`)
participants$`Contact ID` <- as.character(participants$`Contact ID`)
participants$`Address -  Participant ZIP/Postal Code` <- as.character(participants$`Address -  Participant ZIP/Postal Code`)
participants$`Event ID` <- as.character(participants$`Event ID`)
summary(participants)

#change event date from character to date-time
participants$`Event Date` <- mdy_hm(participants$`Event Date`)
class(participants$`Event Date`)
head(participants$`Event Date`, n = 5)
```

```{r}
# acknowledge NA's 
map(participants, ~sum(is.na(.)))
```

## Writing out Jordan's work

```{r}
saveRDS(participants, "C:/Users/Konrad/Desktop/ETM 527/RDS Data/Participants/Participants.rds")
```

