---
title: "Clean and Explore Bike Participants"
author: "Jordan Beary"
date: "January 29, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=TRUE, warning=FALSE}
# load your libraries
library(knitr)
library(stats)
library(readr)
library(dplyr)
library(ggplot2)
library(ggfortify)
library(scales)
library(plotly)
library(plyr)
library(lubridate)
library(tidyverse)
library(reshape2)
library(ggfittext)
library(treemapify)
library(tidyr)
```

## Import data
```{r message=FALSE, warning=FALSE}
participants <- read_csv("Data/Participants/2013-2017 Bike MS Participants.csv")
```

## Basic data exploration
```{r}
class(participants)
dim(participants)
glimpse(participants)
head(participants, n = 5)
tail(participants, n = 5)
summary(participants)
```

Able to see that some fields need to be converted from integers to characters. Also, easily able to identify which fields will need to be rid of NA values. 

```{r}
# need to change fields from int to char 
participants$`Team ID` <- as.character(participants$`Team ID`)
class(participants$`Team ID`)
participants$`Contact ID` <- as.character(participants$`Contact ID`)
participants$`Address -  Participant ZIP/Postal Code` <- as.character(participants$`Address -  Participant ZIP/Postal Code`)
participants$`Event ID` <- as.character(participants$`Event ID`)
summary(participants)

#change event date from character to date-time
participants$`Event Date` <- mdy_hm(participants$`Event Date`)
class(participants$`Event Date`)
head(participants$`Event Date`, n = 5)

# change character type to factor for binning
unique(participants$`Is Prior Participant`)
participants$`Is Prior Participant` <- as.factor(participants$`Is Prior Participant`)
levels(participants$`Is Prior Participant`) <- list(No = "N/A", Yes = "Yes")
head(participants$`Is Prior Participant`)

# gender
unique(participants$`Participant Gender`)
participants$`Participant Gender` <- as.factor(participants$`Participant Gender`)
levels(participants$`Participant Gender`) <- list(Male = "Male", Female = "Female")
head(participants$`Participant Gender`)

# team division
unique(participants$`Team Division`)
participants$`Team Division` <- as.factor(participants$`Team Division`)
levels(participants$`Team Division`) <- list(`Friends and Family` = c("Friends and Family", "Family and Friends"), 
                                             Corporate = c("Corporate", "Corporation"), 
                                             Organization = "Organization (Clubs; Civic Groups; etc.)", `
                                             Church Group` = c("Place of Worship", "Religious"), 
                                             School = "School", 
                                             Other = "Other")
head(participants$`Team Division`)

#fiscal year
unique(participants$`Fiscal Year`)
class(participants$`Fiscal Year`)
participants$`Fiscal Year` <- as.character(participants$`Fiscal Year`)
participants$`Fiscal Year` <- as.Date(participants$`Fiscal Year`, "%Y")
participants$`Fiscal Year` <- year(participants$`Fiscal Year`)
head(participants$`Fiscal Year`)
```

```{r}
# acknowledge NA's 
map(participants, ~sum(is.na(.)))
```


```{r}
# some basic data viz
barplot(table(participants$`Participant Occupation`), ylab = "Count", main = "Occupations of our Participants",
        ylim = c(0,2500), las = 2, col=rgb(0.2,0.4,0.6,0.6))
occ.count <- table(participants$`Participant Occupation`)
occ.count

occupation.omitNA <- participants[!is.na(participants$`Participant Occupation`),]

# Occupation Demographic
top20.occupations.viz <- ggplot(data = occupation.omitNA, aes(x = occupation.omitNA$`Participant Occupation`)) + 
  geom_bar(na.rm = TRUE, aes(y = (..count..), fill = `Participant Gender`)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
top20.occupations.viz <- top20.occupations.viz + 
  labs(x = "Occupations") + 
  labs(title = "Frequency of Participant Occupation") + 
  labs(y = "Count")
top20.occupations.viz

# Prior Participant? Yes - No?
prior.par.bar <- barplot(table(participants$`Is Prior Participant`), ylab = "Count", main = "Count of Prior Participants",
        ylim = c(0,40000), las = 2, col=rgb(0.2,0.4,0.6,0.6))

# Team division bar plot
team.division.bar <- barplot(table(participants$`Team Division`), ylab = "Count", main = "Team Division Demographic",
        ylim = c(0,30000), las = 2, col=rgb(0.2,0.4,0.6,0.6))


```

# Let's build our first persona
## Questions to ask and information to gather
### Gender
```{r}
# make a data frame of gender counts to make visualizing easier
gender.df <- as.data.frame(table(participants$`Participant Gender`))
colnames(gender.df) <- c("Gender", "Count")
glimpse(gender.df)
gender.df

# visualize gender counts 
gender.viz <- ggplot(gender.df, aes(x = Gender, y = Count)) + 
  geom_bar(stat = "identity")
gender.viz
```
Overall we see that participants are predominately male. But what about over time? 
```{r}
gender.vip <- data.frame(participants$`Event Date`, participants$`Participant Gender`, participants$`Participant Occupation`,
                          participants$`Team Division`, participants$`Is Prior Participant`, participants$`Address  -  Participant City`,
                          participants$`Total From Participant($)`, participants$`Fiscal Year`, 
                         participants$`Address  -  Participant State/Province`, participants$`Internal Event Name`)
colnames(gender.vip) <- c("Event Date", "Participant Gender", "Participant Occupation", "Team Division", "Is Prior Participant",
                          "Address - Participant City", "Total From Participant($)", "Fiscal Year", "Address - Participant State",
                          "Event Name")

gender.year <- aggregate(gender.vip, by = list(gender.vip$`Participant Gender`, gender.vip$`Fiscal Year`), FUN = length)
colnames(gender.year) <- c("Gender", "Year", "Count")
gender.year[,c(1,2,3)]


g.year.viz <- ggplot(data = gender.year, 
                     aes(x = gender.year$Year, y = gender.year$Count, fill = gender.year$Gender)) +
  geom_bar(stat = "identity")
g.year.viz +
  labs(title = "Annual Participant Gender Count") +
  labs(x = "Event Year") +
  labs(y = "Count")

gt.count <- aggregate(gender.vip$`Participant Gender` ~ gender.vip$`Fiscal Year`, data = gender.vip, FUN = length)
colnames(gt.count) <- c("Year", "Count")
gt.count

gender.time.viz <- ggplot(gender.vip, aes(x = `Event Date`))
```


Who donates more - guys or gals?
```{r warning=FALSE}
# mean donation for each event
gender.gifts <- aggregate(gender.vip, by = list(gender.vip$`Participant Gender`, gender.vip$`Event Date`), FUN = mean)
gender.gifts <-gender.gifts[,c(1,2,9)]
colnames(gender.gifts) <- c("Gender", "Event Date", "Gift Amount")
gender.gifts

g.gift.viz <- ggplot(data = gender.gifts, 
                     aes(x = gender.gifts$`Event Date`, y = gender.gifts$`Gift Amount`, colour = gender.gifts$Gender)) +
  geom_boxplot() + geom_jitter(width = 0.5)
g.gift.viz +
  labs(title = "Mean Participant Donation per Event by Gender") +
  labs(x = "Date") +
  labs(y = "Gift Amount")

# mean donation for each fiscal year
gender.gifts.year <- aggregate(gender.vip, by = list(gender.vip$`Participant Gender`, gender.vip$`Fiscal Year`), FUN = mean)
gender.gifts.year <-gender.gifts.year[,c(1,2,9)]
colnames(gender.gifts.year) <- c("Gender", "Year", "Gift Amount")
gender.gifts.year
  
g.gift.year.viz <- ggplot(data = gender.gifts.year, 
                     aes(x = gender.gifts.year$Year, y = gender.gifts.year$`Gift Amount`, fill = gender.gifts.year$Gender)) +
  geom_bar(stat = "identity")
g.gift.year.viz +
  labs(title = "Mean Participant Donation per Year by Gender") +
  labs(x = "Fiscal Year") +
  labs(y = "Gift Amount")
  
```
# Where are the participants coming from?
```{r}
# for city
city.count <- as.data.frame(table(c(participants$`Address  -  Participant City`)))
colnames(city.count) <- c("City", "Count")

top.city <- head(arrange(city.count,desc(city.count$Count)), n = 50)

tree.city <- ggplot(top.city, aes(area = Count, fill = Count, label = City)) +
  geom_treemap()+
  geom_treemap_text(fontface = "italic", colour = "white", place = "centre",
                    grow = TRUE)
tree.city

# for state
state.count <- as.data.frame(table(c(participants$`Address  -  Participant State/Province`)))
colnames(state.count) <- c("State", "Count")

top.state <- head(arrange(state.count,desc(state.count$Count)), n = 10)

tree.state <- ggplot(top.state, aes(area = Count, fill = Count, label = State)) +
  geom_treemap()+
  geom_treemap_text(fontface = "italic", colour = "white", place = "centre",
                    grow = TRUE)
tree.state
```

# What do they do for work?
```{r}
# top 15 occupations
count.occ <- as.data.frame(table(gender.vip$`Participant Occupation`))
colnames(count.occ) <- c("Occupation", "Count")
top.occ <- head(arrange(count.occ, desc(Count)), n = 15)

# of the top 10 states what are the top occupations
state.occ <- as.data.frame(
  gender.vip %>%
  select(`Address - Participant State`, `Participant Occupation`) %>%
  filter(`Address - Participant State` == c("CA", "CO", "FL", "CT", "DE", "AL", "AZ", "VA", "PA", "TX")) %>%
  group_by(`Participant Occupation`, `Address - Participant State`) %>%
  tally(!is.na(`Participant Occupation`)) %>%
  top_n(20)
)
state.occ$`Address - Participant State` <- as.character(state.occ$`Address - Participant State`)
spread.occ <- spread(state.occ, `Address - Participant State`, n) # this works but check the data ...very sus

spread.occ[is.na(spread.occ)] <- 0
spread.occ <- spread.occ[-50, ]

spread.viz <- ggplot(data = spread.occ, aes(x = `Participant Occupation`)) + 
  geom_bar()
spread.viz

# which occupation donates the most by state
```


Donations by Event
```{r warning=FALSE}
event.gifts <- aggregate(gender.vip, by = list(gender.vip$`Event Name`, gender.vip$`Fiscal Year`), FUN = mean)
event.gifts <- event.gifts[,c(1,3,9)]
as.data.frame(event.gifts)
colnames(event.gifts) <- c("Event", "Date", "Gift Amount")

event.gifts$Year <- format(event.gifts$Date, "%Y")
event.gifts$Month <- format(event.gifts$Date, "%b")
event.gifts$Day <- format(event.gifts$Date, "%d")
event.gifts$CommonDate <- as.Date(as.character(as.POSIXct(event.gifts$Date)))

event.viz <- ggplot(data = event.gifts, mapping = aes(x = CommonDate, y = `Gift Amount`, shape = Year, colour = Year)) +
    geom_point() +
    geom_line(aes(group = 1))
event.viz <- event.viz + 
  facet_wrap(~Year, ncol = 1, scales = "free_x") + theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  scale_y_continuous() +
  scale_x_date(labels = date_format("%d-%b"), breaks = date_breaks("2 weeks"))
event.viz
```


```{r}
# save RDS 
saveRDS(team.division.bar, "team_division_bar.rds")
saveRDS(occupations.viz, "occupations_vis.rds")
saveRDS(prior.par.bar, "prior_participant.rds")
```


```{r eval=FALSE, include=FALSE}
#occupation.viz <- ggplot(data = occupation.omitNA, aes(x = occupation.omitNA$`Participant Occupation`)) + 
#  geom_bar(aes(y = (..count..)), fill = occupation.omitNA$`Participant Gender`) +
#  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
#occupation.viz + 
#  labs(x = "Occupations") + 
#  labs(title = "Frequency of Participant Occupation") + 
#  labs(y = "Count")

#occupation.viz <- ggplot(data = subset(participants, `Participant Occupation` != "NA"), aes(x = participants$`Participant Occupation`)) + 
#  geom_bar(na.rm = TRUE, aes(y = (..count..), fill = `Participant Gender`)) +
#  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
#occupation.viz + labs(x = "Occupations") + 
#  labs(title = "Frequency of Participant Occupation") + 
#  labs(y = "Count")
```


```{r}
donations.ramya <- read_csv("C:\\Users\\J-Blazer\\OneDrive\\ETM 527 DataSets\\Donations\\2017 Bike Donations.csv")
occupations.ramya <- merge(donations.ramya[, c("Event_ID", "Gift Amount")])

funds <- donations.ramya %>%
 group_by(Participant Occupation)%>%
 summarise(Gift.Amount)
```

